
---
title: "Analysis on eQTL results w/wo transcriptome diversity as covariate using Keele's data and eqtl method"
output:
    html_document:
        fig_crop: no
        toc: true
        toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
options(digits=4)
```

```{r libraries}
source('../../R/ggthemes.R', chdir=T)
#source('../../R/gtex.R', chdir=T)
source('../../R/misc.R', chdir=T)
#source('../../R/GWAS.R', chdir=T)
#source('../../R/idConverter.R', chdir=T)
source('../../R/plots.R', chdir=T)
library('readr')
library('ggplot2')
#library('purrr')
library('dplyr')
library('tidyr')
library('kableExtra')
library('patchwork')
#library('broom')
#library('forcats')
#library('broom')
library('stringr')

```

```{r opts}
# Colour of points and bars
FILL_COL3 <- 'darkslategray'
FILL_COL2 <- 'cadetblue4'
FILL_COL1 <- 'cadetblue2'

# Colour for del/amp/totaa
FILL_COL4 <- 'forestgreen'
FILL_COL5 <- 'coral'
FILL_COL6 <- 'dodgerblue3'


# Files
eqtl_trans_folder <- '/scratch/users/paedugar/cnv_gtex_project/transcriptome_diversity_other_datasets/keele_paper_eqtls_covariate_transcriptome_variability/'
eqtl_trans_batch_folder <- '/scratch/users/paedugar/cnv_gtex_project/transcriptome_diversity_other_datasets/keele_paper_eqtls_covariate_transcriptome_variability_and_batch/'
eqtl_batch_folder <- '/scratch/users/paedugar/cnv_gtex_project/transcriptome_diversity_other_datasets/keele_paper_eqtls_covariate_batch/'
eqtl_no_covs_folder <- '/scratch/users/paedugar/cnv_gtex_project/transcriptome_diversity_other_datasets/keele_paper_eqtls_no_covariate/'
```

```{r ref.label='source'}
```


# Goal
This analysis aims to understand to what extend transcriptome diverisity can influence eQTL analysis


# Results

Load the data

```{r load_data, results='hide'}
# eqtl results using transcriptome diversity as covariate
eqtl_trans <- read_eqtl(eqtl_trans_folder, 'transcriptome_diversity_as_covariates')
eqtl_trans_batch <- read_eqtl(eqtl_trans_batch_folder, 'transcriptome_diversity_batch_as_covariates')
eqtl_batch <- read_eqtl(eqtl_batch_folder, 'batch_as_covariates')
eqtl_no_covs <- read_eqtl(eqtl_no_covs_folder, 'no_covariates')

eqtl_results <- bind_rows(eqtl_trans, eqtl_trans_batch, eqtl_batch, eqtl_no_covs)

eqtl_results_merged <- eqtl_results %>%
    group_by(group, id_names) %>%
    mutate(fdr=p.adjust(nominal_pvalue, method='BH')) %>%
    ungroup()

# Load expression associations with transcriptome diversity
#expression_association_files <- list.files(expression_association_dir, recursive=T, pattern='significant.txt.gz$', full.names=T)
#expression_associations <- concatenate_table_files2(expression_association_files, id_names=basename(dirname(expression_association_files)))
#
#eqtl_results_merged <- left_join(eqtl_results, select(expression_associations, c(FDR, id_names, gene))) 
#eqtl_results_merged$expression_association_status <- ifelse(is.na(eqtl_results_merged$FDR), 'not_associated_w_transDiv', 'associated_w_transDiv')
```



Plot distribution of pvalues in both analysis
```{r plot_pvalues, fig.width=15}
ggplot(eqtl_results_merged, aes(x=nominal_pvalue)) +
    geom_histogram(fill='grey60', colour='grey20', bins=100) +
    facet_grid(id_names~group, scales='free_y') +
    theme_sleek()
```

QQ plot like batch as covariates vs no covariates
```{r fig.width=12}

toPlot <- eqtl_results_merged %>%
    dplyr::select(gene_id, id_names, group, nominal_pvalue) %>%
    pivot_wider(names_from=group, values_from=nominal_pvalue)


ggplot(toPlot, aes(x=-log10(no_covariates), y=-log10(batch_as_covariates))) +
    geom_point(alpha=0.5,size=0.8) +
    geom_abline(slope=1, intercept=0, colour=FILL_COL2) +
    facet_grid(.~id_names) +
    #xlim(0,50) +
    #ylim(0,50) +
    theme_sleek()
```


QQ plot like transcriptome diversity vs no covariates
```{r fig.width=12}

toPlot <- eqtl_results_merged %>%
    dplyr::select(gene_id, id_names, group, nominal_pvalue) %>%
    pivot_wider(names_from=group, values_from=nominal_pvalue)


ggplot(toPlot, aes(x=-log10(no_covariates), y=-log10(transcriptome_diversity_as_covariates))) +
    geom_point(alpha=0.5,size=0.8) +
    geom_abline(slope=1, intercept=0, colour=FILL_COL2) +
    facet_grid(.~id_names) +
    #xlim(0,50) +
    #ylim(0,50) +
    theme_sleek()
```

QQ plot like batch vs transcriptome diversity
```{r fig.width=12}

toPlot <- eqtl_results_merged %>%
    dplyr::select(gene_id, id_names, group, nominal_pvalue) %>%
    pivot_wider(names_from=group, values_from=nominal_pvalue)


ggplot(toPlot, aes(x=-log10(batch_as_covariates), y=-log10(transcriptome_diversity_as_covariates))) +
    geom_point(alpha=0.5,size=0.8) +
    geom_abline(slope=1, intercept=0, colour=FILL_COL2) +
    facet_grid(.~id_names) +
    #xlim(0,50) +
    #ylim(0,50) +
    theme_sleek()
```

QQ plot like batch vs batch + transcriptome diversity
```{r fig.width=12}

toPlot <- eqtl_results_merged %>%
    dplyr::select(gene_id, id_names, group, nominal_pvalue) %>%
    pivot_wider(names_from=group, values_from=nominal_pvalue)


ggplot(toPlot, aes(y=-log10(transcriptome_diversity_batch_as_covariates), x=-log10(batch_as_covariates))) +
    geom_point(alpha=0.5,size=0.8) +
    geom_abline(slope=1, intercept=0, colour=FILL_COL2) +
    facet_grid(.~id_names) +
    #xlim(0,50) +
    #ylim(0,50) +
    theme_sleek()
```






```{r source}
read_eqtl <- function(eqtl_dir, group) {
    
    cor_files <- list.files(eqtl_dir, full.names=T, recursive=T, pattern='txt$')
    cor_ids <- str_split_fixed(basename(cor_files), '\\.', n=2)[,1]
    cors <- concatenate_table_files2(cor_files, id_names=cor_ids, delim='\t', header=T, read_function=read_delim, progress=F)
    cors$group <- group
    
    return(cors)
}
```

